Interpreting Random Forest Predictions for Firearm Identification Using LIME

### 2018-11-19 paper/Interpreting Random Forest Predictions for Firearm Identification Using LIME.Rmd
added lime options to paper and prepared code to run for a large set of lime options

\begin{quote}
To do this, `lime` draws a sample from a Uniform(0,1) distribution and uses the bin information to compute a sample drawn from the feature distribution.
\end{quote}

@@

  # Specify the options to use with lime 
  hamby224_lime_cases <- data.frame(bin_continuous = rep(TRUE, 18),
                                    quantile_bins = c(rep(TRUE, 9), rep(FALSE, 9)),
                                    nbins = rep(2:10, 2),
                                    use_density = rep(NA, 18)) %>%
  mutate(case = 1:length(nbins)) %>%
  select(case, bin_continuous, nbins, quantile_bins, use_density)

  # Run the lime functions for the specified number of bins
  hamby224_all_results <- pmap(.l = as.list(hamby224_lime_cases[1,-1]),
             .f = run_lime, # run_lime is one of my helper functions
             train = hamby173and252_train %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3)

@@

  # Name the items in the lime list
  names(hamby224_lime) <- map_chr(1:dim(hamby224_lime_cases[1,])[1], function(case) 
      sprintf("case: bin_continuous = %s, quantile_bins = %s, nbins = %0.f, use_density = %s",
              hamby224_lime_cases$bin_continuous[case],
              hamby224_lime_cases$quantile_bins[case],
              hamby224_lime_cases$nbins[case],
              hamby224_lime_cases$use_density[case]))

@@

```{r}
hamby224_lime_comparisons <- hamby224_test_explain %>%
  select(-data, -prediction) %>%
  group_by(case, bin_continuous, quantile_bins, nbins, use_density) %>%
  slice(1) %>%
  ungroup() %>%
  select(bin_continuous, quantile_bins, nbins, use_density, case, set, rf_features, rfscore, model_prediction, model_r2) %>%
  mutate(diff = rfscore - model_prediction,
         mean = (rfscore + model_prediction) / 2)

hamby224_lime_comparisons %>%
  group_by(bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff)^2) / length(diff)) %>%
  arrange(set)

hamby224_lime_comparisons %>%
  filter(nbins == "2", set == "Set 1") %>%
  select(rf_features, diff, -case) %>%
  gather(key = feature, value = feature_value, rf_features) %>%
  select(feature, feature_value, diff) %>%
  ggplot(aes(x = feature_value, y = diff)) + 
  geom_point() + 
  facet_wrap( ~ feature, scales = "free") + 
  geom_hline(yintercept = 0, color = "blue")
```

@@

```{r fig.width = 8, fig.height = 4, eval = FALSE}
ggplot(hamby224_lime_comparisons, aes(x = rfscore, y = model_prediction)) + 
  geom_point(aes(color = model_r2)) + 
  facet_grid(nbins ~ set) + 
  theme_bw() + 
  stat_smooth(method = "lm")
```

@@

```{r fig.width = 8, fig.height = 4, eval = FALSE}
ggplot(hamby224_lime_comparisons, aes(x = rfscore, y = abs(diff))) + 
  geom_point(aes(color = model_r2)) + 
  facet_grid(nbins ~ set) +
  theme_bw()
```

@@

```{r fig.width = 8, fig.height = 4, eval = FALSE}
ggplot(hamby224_lime_comparisons, aes(x = rfscore, y = diff)) + 
  geom_point(aes(color = model_r2)) + 
  facet_grid(nbins ~ set) +
  geom_hline(yintercept = 0, color = "blue") +
  theme_bw()
```

@@

```{r, eval = FALSE}
hamby224_test_explain %>%
  ggplot(aes(x = feature_desc)) + 
  geom_bar() +
  coord_flip() + 
  facet_grid(set ~ nbins)
```

```{r, eval = FALSE}
hamby224_test_explain %>%
  filter(set == "Set 1", nbins == "3") %>%
  mutate(rfscore = round(rfscore, 3)) %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2,
             text = paste('Bullets Compared: ', bullet1, "-", land1, 
                          "vs", bullet2, "-", land2,
                          '\nRandom Forest Score: ', 
                          ifelse(is.na(rfscore), "Missing due to tank rash", rfscore)))) +
  geom_tile(aes(fill = rfscore)) +
  facet_grid(bullet2 ~ bullet1, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "", fill = "RF Score")
```

>>>>

\begin{quote}
To do this, `lime` has several options (mostly quoted from `lime` package for now):
\begin{itemize}
\item bin\_continuous = TRUE should continuous variables be binned?
\item quantile\_bins = TRUE should the ins for n\_bins be based on quantiles or spread evenly
\item n\_bins = 4 number of bins if bin\_continuous is TRUE
\item use\_density = TRUE if bin\_continuous is FALSE, should continuous data be sampled using kernel density estimation (if not, then will assume normal for continuous variable)
\end{itemize}
\end{quote}

@@

  # Specify the options to use with lime 
  hamby224_lime_cases <- data.frame(bin_continuous = c(rep(TRUE, 18), rep(FALSE, 2)),
                                    quantile_bins = c(rep(TRUE, 9), rep(FALSE, 9), rep(NA, 2)),
                                    nbins = c(rep(2:10, 2), rep(NA, 2)),
                                    use_density = c(rep(NA, 18), TRUE, FALSE)) %>%
  mutate(case = 1:length(nbins)) %>%
  select(case, bin_continuous, nbins, quantile_bins, use_density)

  # Run the lime functions for the specified number of bins
  hamby224_all_results <- pmap(.l = as.list(hamby224_lime_cases[,-1]),
             .f = run_lime, # run_lime is one of my helper functions
             train = hamby173and252_train %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3)

@@

  # Name the items in the lime list
  names(hamby224_lime) <- map_chr(1:dim(hamby224_lime_cases)[1], function(case) 
      sprintf("case: bin_continuous = %s, quantile_bins = %s, nbins = %0.f, use_density = %s",
              hamby224_lime_cases$bin_continuous[case],
              hamby224_lime_cases$quantile_bins[case],
              hamby224_lime_cases$nbins[case],
              hamby224_lime_cases$use_density[case]))

@@

```{r}
# Create the lime comparison data if the file does not already exist
if(!file.exists("../data/hamby224_lime_comparisons.rds")) {

  # Create a data frame with the interesting information relating to the differe
  # evaluations of lime and compute the difference and mean between the rf and rr
  # model predictions
  hamby224_lime_comparisons <- hamby224_test_explain %>%
    select(-data, -prediction) %>%
    group_by(case, bin_continuous, quantile_bins, nbins, use_density) %>%
    slice(1) %>%
    ungroup() %>%
    select(bin_continuous, quantile_bins, nbins, use_density, 
           set, case, rf_features, rfscore, model_prediction, model_r2) %>%
    mutate(diff = rfscore - model_prediction,
           mean = (rfscore + model_prediction) / 2)

  # Save the lime comparison data frame
  saveRDS(hamby224_lime_comparisons, "../data/hamby224_lime_comparisons.rds")

  # Load in the lime comparison data frame
  hamby224_lime_comparisons <- readRDS("../data/hamby224_lime_comparisons.rds")

```{r}
hamby224_lime_comparisons %>%
  group_by(bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff)^2) / length(diff)) %>%
  arrange(set)
```

```{r}
hamby224_lime_comparisons %>%
  filter(nbins == "2", set == "Set 1") %>%
  select(rf_features, diff, -case) %>%
  gather(key = feature, value = feature_value, rf_features) %>%
  select(feature, feature_value, diff) %>%
  ggplot(aes(x = feature_value, y = diff)) + 
  geom_point() + 
  facet_wrap( ~ feature, scales = "free") + 
  geom_hline(yintercept = 0, color = "blue")
```

@@

```{r fig.width = 8, fig.height = 4}
ggplot(hamby224_lime_comparisons, aes(x = rfscore, y = model_prediction)) + 
  geom_point(aes(color = model_r2)) + 
  facet_grid(nbins ~ set) + 
  theme_bw() + 
  stat_smooth(method = "lm")
```

@@

```{r fig.width = 8, fig.height = 4}
ggplot(hamby224_lime_comparisons, aes(x = rfscore, y = abs(diff))) + 
  geom_point(aes(color = model_r2)) + 
  facet_grid(nbins ~ set) +
  theme_bw()
```

@@

```{r fig.width = 8, fig.height = 4}
ggplot(hamby224_lime_comparisons, aes(x = rfscore, y = diff)) + 
  geom_point(aes(color = model_r2)) + 
  facet_grid(nbins ~ set) +
  geom_hline(yintercept = 0, color = "blue") +
  theme_bw()
```

@@

```{r}
hamby224_test_explain %>%
  ggplot(aes(x = feature_desc)) + 
  geom_bar() +
  coord_flip() + 
  facet_grid(set ~ nbins)
```

```{r}
hamby224_test_explain %>%
  filter(set == "Set 1", nbins == "2") %>%
  mutate(rfscore = round(rfscore, 3)) %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2,
             text = paste('Bullets Compared: ', bullet1, "-", land1, 
                          "vs", bullet2, "-", land2,
                          '\nRandom Forest Score: ', 
                          ifelse(is.na(rfscore), "Missing due to tank rash", rfscore)))) +
  geom_tile(aes(fill = rfscore)) +
  facet_grid(bullet2 ~ bullet1, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "", fill = "RF Score")
```
